---
title: "Homework assignment 1"
author: "Malaikah Ahmad"
format: pdf
---
## Question 1
```{r}
library(babynames)
library(dplyr)
```

## Question 2

## There are 1924665 observations of 5 variables


## Question 3
```{r}
library(tibble)
data_dictionary <- tibble::tibble(
Variable_name = c ("year", "sex", "name", "n", "prop"),
Data_Type = c ("integer", "factor", "character", "integer", "numeric"),
Description = c (
"Year in which the baby was born",
"Sex of the baby (M = Male, F = Female)",
"Name of the baby",
"Number of babies with that name born in a given year",
"Proportion of babies with that name out of all babies of the same sex born in that year"
)
)

print(data_dictionary)
```

## Question 4
```{r}
range(babynames$year)

```

## The range of years covered in babynames is 1880 to 2017

## Question 5
```{r}
babynames_no_n <- dplyr::select(babynames, year, sex, name, prop)

```

## Question 6

## One reason for keeping the variable prop is to focus on the relative popularity of a given name, independent of changes in the total population. The variable n is not useful in this case. 

## Question 7a
```{r}
second_millenium <- babynames_no_n %>% 
filter (year <= 2000) %>% 
group_by(sex) %>% 
slice_max(prop, n = 1) %>% 
ungroup() 

second_millenium
```

## The most popular names for the 2nd millenium are Mary and John

## Question 7b
```{r}
third_millenium <- babynames_no_n %>% 
filter (year >= 2001) %>% 
group_by(sex) %>% 
slice_max(prop, n = 1) %>% 
ungroup()

third_millenium
```

## The most popular names for the 3rd millenium are Emily and Jacob

## Question 8
```{r}
popular_names <- babynames %>% 
filter (year >=2000,  year <=2012, grepl("^[QVX]", name)) %>% 
group_by (name) %>% 
summarize (total_n = sum(n)) %>% 
arrange(desc(total_n))

popular_names
```

## The most popular names beginning with the letter Q, V, and X between the years of 2000-2012 are Quinn, Victoria, and Xavier

## Question 9
```{r}
babynames_decade <- babynames %>% 
mutate(decade = floor(year/10)*10)
```

## Question 10
```{r}
mean_median_babies <- babynames_decade %>%
  group_by(decade, sex) %>%
  summarize(
    mean_n = mean(prop * n(), na.rm = TRUE),
    median_n = median(prop * n(), na.rm = TRUE)
  ) %>%
  ungroup()
  mean_median_babies
```

## Question 11

```{r}
popularMalaikah <- babynames_decade |>
  filter(name == "Malaikah") |>
  arrange (desc(n)) |>
  slice(1)

popularSupervisor <- babynames_decade |>
  filter(name == "Aimee") |>
  arrange (desc(n)) |>
  slice(1)

popularJack <- babynames_decade |>
  filter(name == "Jack") |>
  arrange(desc(n)) |>
  slice(1)

popularScott <- babynames_decade |>
  filter(name == "Scott") |>
  arrange (desc(n)) |>
  slice (1)
```

## My name (Malaikah) was most popular in the 2000s, specifically 2002.
## My Supervisor's name (Aimee) was most popular in the 1970s, specifically 1977. 
## The name Jack was most popular in the 1920s, specifically 1927.
## The name Scott was most popular in the 1970s, specifically 1971.
